Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists

### 2024-02-29 paper.Rmd
paper update

---
title: "Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists"
author:
  - name: Ronald M. Visser
    orcid: 0000-0001-6966-1729
    affil-id: 1
  - name: Annemarie Jutte
    orcid: 0000-0000-0000-0000
    affil-id: 1
  - name: Doug Rocks-Macqueen 
    orcid: 0000-0000-0000-0000
    affil-id: 2
  - name: Iza Romanowska 
    orcid: 0000-0000-0000-0000
    affil-id: 3
  - name: Tom Brughmans
    orcid: 0000-0000-0000-0000
    affil-id: 3
  - name: Laura van der Knaap 
    orcid: 0000-0000-0000-0000
    affil-id: 4
  - name: Karsten Lambers 
    orcid: 0000-0000-0000-0000
    affil-id: 4
  - name: Kenneth Aitchison 
    orcid: 0000-0000-0000-0000
    affil-id: 2
affiliations:
  - id: 1
    name: Saxion University of Applied Sciences.
  - id: 2
    name: Landward Research Teoranta.
  - id: 3
    name: Aarhus University.
  - id: 4
    name: Leiden University.
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry:
  - margin=1in
output:
  pdf_document:
    number_sections: yes
    keep_tex: true
    fig_caption: yes
    latex_engine: pdflatex
  html_document:
    number_sections: yes
  word_document:
    number_sections: yes
bibliography: references.bib
csl: journal-of-computer-applications-in-archaeology.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

@@

The surveys were analysed using R [@rcoreteam2023]. The following packages were used for analyses: ggplot2 [@wickham2016], dplyr [@wickham2023a], tidyr [@wickham2023], forcats[@wickham2023b] , lubridate [@grolemund2011] and stringr [@wickham2023c]. The date and the code is available at <https://github.com/RonaldVisser/ABM_tutorials> and (... insert Zenodo reference...)

@@

A large proportion of the participants of the workshops gave us information using the survey before the workshops 172 of (...total number...) participants. The respondents came from many different countries all over the world (see Figure \@ref(fig:nationality). The number of female respondents slightly outnumber the male ones and a small group did not share their gender and two were non-binary \@ref(fig:gender-age). The ages of the respondents ranged from below 20 to over 70, reaching people in different age classes, although the majority was aged between 20 and 40.

@@

Most of the respondents had several skills with computer applications in archaeology (see Figure @ref(fig:computer-skills)), and many of them did have some knowledge of ABM, although a large group was completely new to the subject. The majority had never applied ABM before participating in the workshops with 14 having applied ABM. It is interesting to note that many respondents did know what kind of software was available for ABM (see Figure \@ref(fig:abm-knowledge)).

```{r, computer-skills, fig.cap="The computer skills of the respondents."}
pre_surveys %>%
  select(`Which.computer.skills.do.you.have....Selected.Choice`) %>%
  separate_rows(`Which.computer.skills.do.you.have....Selected.Choice`, sep = ',\\s*') %>%
  filter(`Which.computer.skills.do.you.have....Selected.Choice`!="") %>%
  group_by(`Which.computer.skills.do.you.have....Selected.Choice`) %>%
  ggplot(aes(x=`Which.computer.skills.do.you.have....Selected.Choice`)) + geom_bar() +
    xlab("Computer skills") + coord_flip()

```{r, abm-knowledge, fig.cap="Respondents knowledge and experience with ABM.", fig.show="hold", out.width="50%"}
pre_surveys %>% 
  filter(`Do.you.know.what.Agent.Based.Modeling.is.`!="Not shared") %>% 
  ggplot(aes(x = `Do.you.know.what.Agent.Based.Modeling.is.`)) +
    geom_bar(na.rm = TRUE) + xlab("I know what ABM is") + facet_grid(~Workshop)
pre_surveys %>%
  select(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  filter(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`!="") %>%
  separate_rows(Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice, sep = ',\\s*') %>%
  group_by(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  ggplot(aes(x=`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`)) + geom_bar() +
  xlab("ABM Software known") + coord_flip()

>>>>

---
title: "Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists"
author:
  - name: Ronald M. Visser
    orcid: 0000-0001-6966-1729
    affil-id: 1
  - name: Annemarie Jutte
    orcid: 
    affil-id: 1
  - name: Doug Rocks-Macqueen 
    orcid: 
    affil-id: 2
  - name: Iza Romanowska 
    orcid: 
    affil-id: 3
  - name: Tom Brughmans
    orcid: 
    affil-id: 3
  - name: Laura van der Knaap 
    orcid: 
    affil-id: 4
  - name: Karsten Lambers 
    orcid: 
    affil-id: 4
  - name: Kenneth Aitchison 
    orcid: 
    affil-id: 2
affiliations:
  - id: 1
    name: Saxion University of Applied Sciences.
  - id: 2
    name: Landward Research Teoranta.
  - id: 3
    name: Aarhus University.
  - id: 4
    name: Leiden University.
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry:
  - margin=1in
output:
  bookdown::pdf_document2:
    number_sections: yes
    keep_tex: true
    fig_caption: yes
  html_document:
    number_sections: yes
    fig_caption: yes
  word_document:
    number_sections: yes
bibliography: references.bib
csl: journal-of-computer-applications-in-archaeology.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

@@

The surveys were analysed using R [@rcoreteam2023]. The following packages were used for analyses: ggplot2 [@wickham2016], dplyr [@wickham2023a], tidyr [@wickham2023], forcats[@wickham2023b], lubridate [@grolemund2011] and stringr [@wickham2023c]. The date and the code is available at <https://github.com/RonaldVisser/ABM_tutorials> and (... insert Zenodo reference...)

@@

A large proportion of the participants of the workshops gave us information using the survey before the workshops 172 of (...total number...) participants. The respondents came from many different countries all over the world (see Figure \@ref(fig:nationality). The number of female respondents slightly outnumber the male ones and a small group did not share their gender and two were non-binary (see Figure \@ref(fig:gender-age)). The ages of the respondents ranged from below 20 to over 70, reaching people in different age classes, although the majority was aged between 20 and 40. It seems that the different events also had a slightly different distribution of both gender and age. For example, more older people attended the workshop at the CAA conference in April 2023 and more male people were present during the workshop at the Reuvensdagen in November 2023.

students???

@@

Most of the respondents had several skills with computer applications in archaeology (see Figure \@ref(fig:computer-skills)), and many of them did have some knowledge of ABM, although a large group was completely new to the subject. The majority had never applied ABM before participating in the workshops with 14 having applied ABM. It is interesting to note that many respondents did know what kind of software was available for ABM (see Figure \@ref(fig:abm-knowledge)).

```{r, computer-skills, fig.cap="The computer skills of the respondents.", out.width="50%"}
pre_surveys %>%
  select(`Which.computer.skills.do.you.have....Selected.Choice`) %>%
  separate_rows(`Which.computer.skills.do.you.have....Selected.Choice`, sep = ',\\s*') %>%
  filter(`Which.computer.skills.do.you.have....Selected.Choice`!="") %>%
  group_by(`Which.computer.skills.do.you.have....Selected.Choice`) %>%
  ggplot(aes(x=`Which.computer.skills.do.you.have....Selected.Choice`)) + geom_bar() +
    xlab("Computer skills") + coord_flip()
```

```{r, abm-knowledge, fig.cap="Respondents knowledge and experience with ABM.", fig.show="hold", out.width="50%"}
pre_surveys %>% 
  filter(`Do.you.know.what.Agent.Based.Modeling.is.`!="Not shared") %>% 
  ggplot(aes(x = `Do.you.know.what.Agent.Based.Modeling.is.`)) +
    geom_bar(na.rm = TRUE) + xlab("I know what ABM is") + facet_grid(~Workshop)
pre_surveys %>%
  select(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  filter(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`!="") %>%
  separate_rows(Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice, sep = ',\\s*') %>%
  group_by(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  ggplot(aes(x=`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`)) + geom_bar() +
  xlab("ABM Software known") + coord_flip()
```

### 2024-02-29 paper.tex
paper update

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

The surveys were analysed using R (R Core Team 2023). The following
packages were used for analyses: ggplot2 (Wickham 2016), dplyr (Wickham
et al. 2023), tidyr (Wickham, Vaughan and Girlich 2023), forcats(Wickham
2023a) , lubridate (Grolemund and Wickham 2011) and stringr (Wickham
2023b). The date and the code is available at
\url{https://github.com/RonaldVisser/ABM_tutorials} and (\ldots{} insert
Zenodo reference\ldots)

@@

A large proportion of the participants of the workshops gave us
information using the survey before the workshops 172 of (\ldots total
number\ldots) participants. The respondents came from many different
countries all over the world (see Figure @ref(fig:nationality). The
number of female respondents slightly outnumber the male ones and a
small group did not share their gender and two were non-binary
@ref(fig:gender-age). The ages of the respondents ranged from below 20
to over 70, reaching people in different age classes, although the
majority was aged between 20 and 40.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/nationality-1.pdf}
\caption{The nationality of the participants that filled in the survey.}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/gender-age-1.pdf}
\caption{The gender and age of the respondents for each workshop}
\end{figure}

Most of the respondents had several skills with computer applications in
archaeology (see Figure (\textbf{ref?})(fig:computer-skills)), and many
of them did have some knowledge of ABM, although a large group was
completely new to the subject. The majority had never applied ABM before
participating in the workshops with 14 having applied ABM. It is
interesting to note that many respondents did know what kind of software
was available for ABM (see Figure @ref(fig:abm-knowledge)).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/computer-skills-1.pdf}
\caption{The computer skills of the respondents.}
\end{figure}

>>>>

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

The surveys were analysed using R (R Core Team 2023). The following packages were used for analyses: ggplot2 (Wickham 2016), dplyr (Wickham et al. 2023), tidyr (Wickham, Vaughan and Girlich 2023), forcats(Wickham 2023a), lubridate (Grolemund and Wickham 2011) and stringr (Wickham 2023b). The date and the code is available at \url{https://github.com/RonaldVisser/ABM_tutorials} and (\ldots{} insert Zenodo reference\ldots)

@@

A large proportion of the participants of the workshops gave us information using the survey before the workshops 172 of (\ldots total number\ldots) participants. The respondents came from many different countries all over the world (see Figure \ref{fig:nationality}. The number of female respondents slightly outnumber the male ones and a small group did not share their gender and two were non-binary (see Figure \ref{fig:gender-age}). The ages of the respondents ranged from below 20 to over 70, reaching people in different age classes, although the majority was aged between 20 and 40.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/nationality-1.pdf}
\caption{\label{fig:nationality}The nationality of the participants that filled in the survey.}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/gender-age-1.pdf}
\caption{\label{fig:gender-age}The gender and age of the respondents for each workshop}
\end{figure}

Most of the respondents had several skills with computer applications in archaeology (see Figure \ref{fig:computer-skills}), and many of them did have some knowledge of ABM, although a large group was completely new to the subject. The majority had never applied ABM before participating in the workshops with 14 having applied ABM. It is interesting to note that many respondents did know what kind of software was available for ABM (see Figure \ref{fig:abm-knowledge}).

\begin{figure}
\includegraphics[width=0.5\linewidth]{paper_files/figure-latex/computer-skills-1} \caption{The computer skills of the respondents.}\label{fig:computer-skills}
\end{figure}
